---
title: "Final Project (Happiness)"
author: "Jullian Alcantaraa"
date: "2024-04-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(carData)
```


```{r}
happy <- read.csv("WHR2023.csv")
attach(happy)
head(happy)
dim(happy)
str(happy)
n1 <- nrow(happy)
summary(happy)
```


```{r}
happy2 <- happy[,- c(12,4,5)]
colnames(happy2)
```


```{r}
pairs(Ladder.score ~  upperwhisker + lowerwhisker + Healthy.life.expectancy + Perceptions.of.corruption, data = happy2)
```

```{r}
pairs(Ladder.score ~  Explained.by..Social.support  + Explained.by..Healthy.life.expectancy + Explained.by..Freedom.to.make.life.choices + Dystopia...residual, data = happy2)
```




```{r}
mod <- lm(Ladder.score ~ . - Country.name, data = happy2)
summary(mod)

mod1 <- lm(Ladder.score ~ 1 , data = happy2)
summary(mod1)
```

Backward AIC and BIC
```{r}
backstepa <- step(mod, direction = "backward", trace = 0) # AIC (10 Predictors)
backstepa
backstepb <- step(mod, direction = "backward", trace = 0, k = log(n1)) #BIC (9 Predictors)
backstepb
```

Forward AIC and BIC
```{r}
stepa <- step(mod, scope = list(lower = mod1, upper = mod), direction = "forward", trace = 0) # AIC
stepa
stepb <- step(mod, scope = list(lower = mod1, upper = mod), direction = "forward", trace = 0, k = log(n1)) #BIC
stepb
```


```{r}
anova(backstepa, backstepb, stepa, stepb)
```



Refitting the new model
```{r}
fit <- lm(Ladder.score ~ upperwhisker + lowerwhisker + Logged.GDP.per.capita + 
    Healthy.life.expectancy + Generosity + Perceptions.of.corruption + 
    Explained.by..Social.support + Explained.by..Healthy.life.expectancy + 
    Explained.by..Freedom.to.make.life.choices + Dystopia...residual, 
    data = happy)
summary(fit)
plot(fit, 1)
plot(fit,2)
shapiro.test(residuals(fit))
```


Removing the Upper, Lower Whisker and Explained by Healthy Life expectancy predictors also (Final Model)
```{r}
library(car)
fit2 <- lm(Ladder.score ~  Logged.GDP.per.capita +
    Healthy.life.expectancy + Generosity + Perceptions.of.corruption + 
    Explained.by..Social.support  + 
    Explained.by..Freedom.to.make.life.choices + Dystopia...residual, data = happy2)
summary(fit2)
plot(fit2, 1)
plot(fit2,2)
plot(fit2, 4)
shapiro.test(residuals(fit2))
```
Confidence Interval
```{r}
confint(fit2)
```

Coefficients
```{r}
coefficients(fit2)
```

```{r}
avPlots(fit2)
```


```{r}
library(faraway)
library(psych)
round(cor(happy2[, 2:8]), 2)
cor.plot(happy2[, 2:8])
```

$\widehat{\text{Ladder Score}}= -2.58 +0.36\text{GDP per Capita} + 0.027\text{Healthy Life Expectancy} + 0.54\text{Generosity} - 0.71 \text{Perception of Corruption} + 1 \text{Social Support} + 1 \text{Freedom} + 1 \text{Dystopian Residual}.$

VIF
```{r}
vif(fit2)
```



```{r}
detach(happy)
```

